---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![Travis-CI Build Status](https://travis-ci.org/SMAC-Group/exts.svg?branch=master)](https://travis-ci.org/SMAC-Group/exts)

# Experimental Features for Time Series (`exts`)

The package as it presently stands contains functions experimental
and may break at a moments notice. These functions are being developed for use
within Fall 2016 courses at the University of Illinois at Urbana-Champaign (UIUC).
Specifically, students are using these functions within 
STAT 578 (Special Topic): Time Series Forecasting and
STAT 429: Time Series Analysis. Presently, the functions are meant to provide
students with the ability to interact with time series data.
All functions will eventually end up either moving to a different package 
(perhaps `gmwm`)  or may end up being removed. Hence, 
be wary of applying a dependency to `exts` as this is serving as a
very _public_ skunkworks.

# Installation Instructions

Presently, the package can only be obtained through GitHub. Plans for a 
CRAN-based release are tentative. At some point, this may be added to SMAC's
package mirror for the install.

## Installing the package through GitHub (Developmental)

For users who are interested in having the latest and greatest developments
within time series, this option is ideal. Though, there is considerably more
work that a user must do to have a stable version of the package. 
**The setup to obtain the development version is platform dependent.**

Specifically, one **must** have a compiler installed on your system that is 
compatible with R.

For help on obtaining a compiler consult:

* [OS X](http://thecoatlessprofessor.com/programming/r-compiler-tools-for-rcpp-on-os-x/)
* [Windows](http://stat385.thecoatlessprofessor.com/announcements/2016/07/25/install-rtools-for-rcpp/)

Generally speaking, Linux users should have a compiler that is compatible with
R already installed on their system.

With the system dependency taken care of, we continue on by installing the R specific package dependencies and finally the package itself by doing the following in an R session: 

```{r, eval = F}
# Install dependencies
# install.packages("devtools")

# Install the package from GitHub without Vignettes/User Guides
devtools::install_github("SMAC-Group/exts")

# Install the package from GitHub with Vignettes/User Guides
# Note: This will be a longer install as the vignettes must be built.
devtools::install_github("SMAC-Group/exts", build_vignettes = TRUE)
```

## Using the package

To use the package simply load it into R via:

```{r load_library, message = FALSE}
library("exts")
```

Presently, `gmwm` and `magrittr` will autoload allowing you to access
`ts.model`, `gts`, and the pipe (`%>%`) operator.

# Warning

Any function or feature you see here is likely to change or be integrated with
a different package.
