---
title: "Experimental Time Series Features"
author: "JJB"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Experimental Time Series Features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, fig.width = 7)
```

# Introduction

The package as it presently stands contains functions that are being used
within Fall 2016 courses at the University of Illinois at Urbana-Champaign (UIUC).
Specifically, students are using these functions within 
STAT 578 (Special Topic): Time Series Forecasting and
STAT 429: Time Series Analysis. Presently, the functions are meant to provide
students with the ability to interact with time series data. All functions
will eventually end up either moving to a different package (perhaps `gmwm`) 
or may end up being removed.

## Pipable Time Series Functions

Outside of being used as an educational and practical tool, the package also
serves to provide tenets of pipable data within a time series context. Thus, the
package is primarily designed with the notion of the piping operator (`%>%`)
fully in play as well as other _R_ related tools like `dplyr` and `tidyr`.
As this package is experimental, the implementations may shift significantly.

# Installation Instructions

Presently, the package can only be obtained through GitHub. Plans for a 
CRAN-based release are tentative. At some point, this may be added to SMAC's
package mirror for the install.

## Installing the package through GitHub (Developmental)

For users who are interested in having the latest and greatest developments
within time series, this option is ideal. Though, there is considerably more
work that a user must do to have a stable version of the package. 
**The setup to obtain the development version is platform dependent.**

Specifically, one **must** have a compiler installed on your system that is 
compatible with R.

For help on obtaining a compiler consult:

* [OS X](http://thecoatlessprofessor.com/programming/r-compiler-tools-for-rcpp-on-os-x/)
* [Windows](http://stat385.thecoatlessprofessor.com/announcements/2016/07/25/install-rtools-for-rcpp/)

Generally speaking, Linux users should have a compiler that is compatible with
R already installed on their system.

With the system dependency taken care of, we continue on by installing the R specific package dependencies and finally the package itself by doing the following in an R session: 

```{r, eval = F}
# Install dependencies
# install.packages("devtools")

# Install the package from GitHub without Vignettes/User Guides
devtools::install_github("SMAC-Group/exts")

# Install the package from GitHub with Vignettes/User Guides
# Note: This will be a longer install as the vignettes must be built.
devtools::install_github("SMAC-Group/exts", build_vignettes = TRUE)
```

## Using the package

To use the package simply load it into R via:

```{r load_library, message = FALSE}
library("exts")
```

Presently, `gmwm` and `magrittr` will autoload allowing you to access
`ts.model`, `gts`, and the pipe (`%>%`) operator.


# Empirical and Theoretical (R)ACF and PACF

One of the key features of time series resolves around understanding the
dependency that exists between observations using autocorrelation.
There exists four built in functions that enable the computation of the
(Robust) Autocorrelation Function [(R)ACF] and Partial Autocorrelation Function
(PACF). These functions come in two flavors: 

1. empirical or sample-based via `emp_`
2. theoretical ARMA processes via `theo_`

The prior function handles . 
The later is restricted to only analyzing autoregressive moving average processes
through the `ARMA()` object found in `gmwm`. 

To tell what kind of function was graphed, the graph heading will contain
either "Empirical" or "Theoretical" along side the type of computation
(e.g. (R)ACF or PACF).

## Empirical (`emp_`) (R)ACF and PACF

Empirical graphs for (R)ACF and PACF are great for analyzing sample time series
data to determine the dependency structure. 

```{r acf_pacf}
# Set seed for reproducibility
set.seed(8672)

# Generate an AR1
xt = gen.gts(AR1(phi = 0.6, sigma2 = 1))

# Obtain the empirical ACF
vals = emp_acf(xt)
autoplot(vals)

# Obtain the empirical RACF (Robust ACF)
vals = emp_acf(xt, robust = TRUE)
autoplot(vals)

# Obtain the covariance
vals = emp_acf(xt, type = "covariance")
autoplot(vals)

# Obtain the empirical PACF
vals = emp_pacf(xt)
autoplot(vals)

# Obtain both the empirical ACF & PACF
vals = emp_corr(xt)
autoplot(vals)
```

## Theoretical (`theo_`) ACF and PACF graphs for ARMA Models

Theoretical graphs for ACF and PACF are also available separately via 
`theo_acf()` and `theo_pacf()` or combined into one via `theo_corr()`.

```{r theo_arma_plots}
model = ARMA(ar = c(.50, -0.25), ma = .20)

# Theoretical ACF
vals = theo_acf(model)
autoplot(vals)

# Theoretical PACF
vals = theo_pacf(model)
autoplot(vals)

# Create a theoretical graph
vals = theo_corr(model)
autoplot(vals)
```

# Exploratory Data Analysis [EDA] (`eda_()`)

Exploratory Data Analysis allows for visually and quantitatively obtaining
graphing information.

## Observe ACF and PACF Graphs

Often times, it is helpful to observe the time series, ACF, and PACF graphs 
together. To do so, use:

```{r eda_ts, fig.width = 8}
# Bring in time series data
xt = gts(log(lynx), start = 1821, freq = 1)

# Obtain Empirical ACF & PACF alongside data
vals = eda_ts(xt)

# Graph Data, Empirical (R)ACF, and Empirical PACF
autoplot(vals)
```

## Compare Classical and Robust ACF

To observe difference between the classical and robust ACF, one can use `compare_acf()`

```{r compare_robust, fig.width = 8}
# Bring in time series data
xt = gts(log(lynx), start = 1821, freq = 1)

# Obtain the ACF values
vals = compare_acf(xt)

# Graph data
autoplot(vals)
```

## Data Transform Previews

Often times, it is helpful to transform the data to attempt to make the time 
series appear stationary. 

```{r eda_change, fig.height = 6}
# Default transforms
vals = eda_change(gnp)
autoplot(vals)

# Default transforms but applied together in the second case
vals = eda_change(gnp, both = TRUE)
autoplot(vals)

# Different Transforms and applied separately in the second case
vals = eda_change(gnp, t1 = sqrt, t2 = log10)
autoplot(vals)
```

# Diagnostic Functions

The style for diagnostic features has all functions being prefixed by `diag_`.
Each function has its own graphing utility that can be called with either
`autoplot` or `plot` as before.

## Residual extraction and standardization (`diag_resid`)

Residuals are able to be extract from `arima` and `lm` classes in addition to
manipulated to be standardized via `resid / sd(resid)` through a `std = TRUE`
flag.

```{r residual_plots}
# Extract and standardize residuals
model = arima(sunspot.year, c(9,0,0))
resids = diag_resid(model, std = TRUE)

# Graph Histogram
autoplot(resids, type = "hist")

# Graph Histogram
autoplot(resids, type = "resid")

# Graph Histogram
autoplot(resids, type = "both")
```

## Quantile-Quantile Normal Plot (`diag_qq`)

The `diag_qq` plot creates a quantil-quantile plot that contains a fitted
normal line. Typically, this plot is used to evaluate the residuals of a model.

```{r qq_plot}
# Observing residuals of an Arima process
model = arima(sunspot.year, c(9,0,0))
vals = diag_qq(model, std = TRUE)
autoplot(vals)

# Observe time series
vals = diag_qq(sunspot.year, std = TRUE)
autoplot(vals)
```

## Wavelet Variance (WV) Estimate of White Noise (`diag_wv`)

Wavelet Variance provides an estimation technique to display whether the 
observed process is a White Noise. If the orange line follows the dark blue 
line closely or remains within the blue confidence interval, then the process
is likely to be a white noise.

```{r wv_plot}
# White Noise Process
model = arima(sunspot.year, c(9,0,0))
vals = diag_wv(model)
autoplot(vals)
```

## Portmanteau Tests

The package provides a convient wrapper into _R_'s `Box.test()` function that
separates the different tests, 'Ljung-Box' and 'Box-Pierce', to assess the null
hypothesis of independence in a given time series. The function makes available
a `data.frame` containing: `lag`, `p_value`, `statistic` for each of the tests
run that can then be used to achieve a diagnostic plot like so:

```{r portmanteau_test}
# Ljung-Box test
model = arima(sunspot.year, c(9,0,0))
vals = diag_ljungbox(model)
autoplot(vals)

# Box-Pierce Test
vals = diag_boxpierce(model)
autoplot(vals)
```

## Residual vs. Fitted Values (Developmental)

**Warning: The inputs for this function will change.**

One further useful plot is to compare the residuals to the fitted values.

```{r fitted_residual}
xt = sunspot.year
model = arima(xt, c(9,0,0))

# Note additional manipulation is required outside the function
# to obtain the two parameters are required
resids = resid(model)
fit = xt - resids

vals = diag_fitted(fit, resids)
autoplot(vals)
```

## Master Diagnostic Plot (Developmental)


**Warning: This function will change in the future to adhere more to a chained
argument**

Each of the diagnostic plots are available standalone. However, there is benefits
to plotting all of the diagnostic information simultaneously. Hence, the `diag_ts()`
function that will calculate all of the above values and provide a diagnostic plot.

```{r diag_ts, fig.height = 5, fig.width = 8}
# Observing residuals of an Arima process
xt = sunspot.year
model = arima(xt, c(9,0,0))

# Note two parameters are required
vals = diag_ts(model, xt)
autoplot(vals)
```

# Selecting an appropriate time series model

There are many ways to go about selecting an ideal time series model. Within
the package there are `select_arima()`, `select_arma()`, `select_ar()`, 
and `select_ma()` functions that build all of the time series models alongside
multiple Information Criterion (IC) such as:

- Akaike information criterion (AIC): $AIC = - 2{L_{\max }} + 2k$
- Bayesian information criterion (BIC): $BIC = - 2{L_{\max }} + klog(n)$
- Hannan–Quinn information (HQC): $HQC =  - 2{L_{\max }} + 2k\log \left( {\log \left( n \right)} \right)$. 

where $L_{\max }$ is the log-likelihood, $k$ is the number of parameters, 
and $n$ is sample size.

To acess the best model according to one of the criteria, use: `best_model(x, ic="aic")`

## Selecting one dimension `select_ar()` and `select_ma()`

```{r}
# Bring in time series data
xt = gts(diff(log(gnp)), start = 1700, freq = 1)
model_info = select_ar(xt, p.min = 1, p.max = 8)

# Obtain a plot of the model selection criteria
autoplot(model_info)

# Obtain the best model (AIC is default)
best_model(model_info)

# Obtain the best model with BIC
bmodel = best_model(model_info, ic = "bic")

bmodel

# See the ACF of the residuals for the model
vals = emp_corr(bmodel)
autoplot(vals)
```

## Selecting multiple dimensions `select_arima()` and `select_arma()`

**Warning: These procedures have been known to have issues with the initial
range of parameters to attempt. You may wish to change default `*.min` and `*.max`
when applicable. At a later time, support for functional errors should resolve
this issue.**

```{r}
# Bring in time series data
xt = gts(diff(log(gnp)), start = 1700, freq = 1)
model_info = select_arma(xt, p.max = 4, q.max = 3)

# Obtain a plot of the model selection criteria
autoplot(model_info)

# Obtain the best model (AIC is default)
best_model(model_info)

# Obtain the best model with BIC
bmodel = best_model(model_info, ic = "bic")

bmodel

# See the ACF of the residuals for the model
vals = emp_corr(bmodel)
autoplot(vals)
```

# Forecasting Abilities



## Other Packages

The textbook "Time Series Analysis and Its Applications: With Examples in R" (ASTSA)
provides the `astsa` package. This package is able to provide better forecasting
when it comes to model fitting temporarily. The code below is found on page
165 of the ASTSA text.

```{r other_package, fig.height = 6, message = FALSE}
library("astsa")
pred = sarima.for(prodn, 12, 2, 1, 1, 0, 1, 3, 12) # forecast
```

# Utilities

The package hosts various utilities to improve working with simulated data.

In particular, the package offers the ability to cast a simulation `matrix` to
`data.frame` for graphing within `ggplot2`. The simulation cast feature is
meant to ease the use of simulated data within `ggplot2`. Thus, support exists
for both casting simulated data stored by row (default) and by column into a
`Round` and `Draw` framework, where `Round` means the simulation iteration and
`Draw` means the statistic observed.

```{r simulation_objects}
# Set Seed
set.seed(5812)

# Generate data
m = matrix(rnorm(10), 2, 5)

# Organize by row
wide_m = cast_simdf(m)

# View first 5 rows
head(wide_m, n = 5)

# Organize by column
long_m = cast_simdf(m, wide = FALSE)

# View first 5 rows
head(long_m, n = 5)

# Graph wide
autoplot(wide_m)

# Graph long
autoplot(long_m)
```
