---
title: "Experimental Time Series Features"
author: "JJB"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Experimental Time Series Features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, fig.width = 7)
```

# Introduction

The package as it presently stands contains functions that are being used
within Fall 2016 courses at the University of Illinois at Urbana-Champaign (UIUC).
Specifically, students are using these functions within 
STAT 578 (Special Topic): Time Series Forecasting and
STAT 429: Time Series Analysis. Presently, the functions are meant to provide
students with the ability to interact with time series data. All functions
will eventually end up either moving to a different package (perhaps `gmwm`) 
or may end up being removed.

To use the package simply load it into R via:

```{r load_library, echo = FALSE, message = FALSE}
library("exts")
```

Presently, `gmwm` and `magrittr` will autoload. 

# Pipable Time Series Functions

Outside of being used as an educational and practical tool, the package also
serves to provide tenets of pipable data within a time series context. Thus, the
package is primarily designed with the notion of the piping operator (`%>%`)
fully in play as well as other _R_ related tools like `dplyr` and `tidyr`.
As this package is experimental, the implementations may shift significantly.

# (R)ACF and PACF

There exists built in functions that enable the computation of the ACF and PACF
graphs. Specifically, these functions are given by:

```{r}
# Set seed for reproducibility
set.seed(8672)

# Generate an AR1
xt = gen.gts(AR1(phi = 0.6, sigma2 = 1))

# Obtain the empirical ACF
xt %>% 
  emp_acf %>%
  autoplot

# Obtain the empirical RACF (Robust ACF)
xt %>% 
  emp_acf(robust = TRUE) %>%
  autoplot

# Obtain the empirical PACF
xt %>% 
  emp_pacf %>%
  autoplot
```

## Observe ACF and PACF Graphs

Often times, it is helpful to observe the time series, ACF, and PACF graphs 
together. To do so, use:

```{r, fig.width = 8}
# Bring in time series data
xt = gts(log(lynx), start = 1821, freq = 1)

xt %>%
  eda_ts %>%
  autoplot
```

## Theoretical ACF and PACF graphs for ARMA Models

Theoretical graphs are also available separately via `theo_acf()` and `theo_pacf()`
or combined into one via `theo_corr()`.

```{r}
model = ARMA(ar = c(.50, -0.25), ma = .20)

# Theoretical ACF
model %>%
  theo_acf %>%
  autoplot

# Theoretical PACF
model %>%
  theo_pacf %>%
  autoplot

# Create a theoretical graph
model %>%
  theo_corr %>%
  autoplot
```

## Compare Classical and Robust ACF

To observe difference between the classical and robust ACF, one can use `compare_acf()`

```{r, fig.width = 8}
# Bring in time series data
xt = gts(log(lynx), start = 1821, freq = 1)

xt %>%
  compare_acf %>%
  autoplot
```


# Diagnostic Functions

The style for diagnostic features has all functions being prefixed by `diag_`.
Each function has its own graphing utility that can be called with either
`autoplot` or `plot` as before.

## Residual extraction and standardization (`diag_resid`)

Residuals are able to be extract from `arima` and `lm` classes in addition to
manipulated to be standardized via `resid / sd(resid)` through a `std = TRUE`
flag.

```{r}
# Extract and standardize residuals
sunspot.year %>%
  arima(c(9,0,0)) %>%
  diag_resid(std = TRUE) -> resids

# Graph Histogram
autoplot(resids, type = "hist")

# Graph Histogram
autoplot(resids, type = "resid")

# Graph Histogram
autoplot(resids, type = "both")
```

## Quantile-Quantile Normal Plot (`diag_qq`)

The `diag_qq` plot creates a quantil-quantile plot that contains a fitted
normal line. Typically, this plot is used to evaluate the residuals of a model.

```{r}
# Observing residuals of an Arima process
sunspot.year %>%
  arima(c(9,0,0)) %>%
  diag_qq(std = TRUE) %>%
  autoplot

# Observe time series
sunspot.year %>%
  diag_qq(std = FALSE) %>%
  autoplot
```

## Wavelet Variance (WV) Estimate of White Noise (`diag_wv`)

Wavelet Variance provides an estimation technique to display whether the 
observed process is a White Noise. If the orange line follows the dark blue 
line closely or remains within the blue confidence interval, then the process
is likely to be a white noise.

```{r}
# White Noise Process
sunspot.year %>%
  arima(c(9,0,0)) %>%
  diag_wv %>%
  autoplot
```

## Portmanteau Tests

The package provides a convient wrapper into _R_'s `Box.test()` function that
separates the different tests, 'Ljung-Box' and 'Box-Pierce', to assess the null
hypothesis of independence in a given time series. The function makes available
a `data.frame` containing: `lag`, `p_value`, `statistic` for each of the tests
run that can then be used to achieve a diagnostic plot like so:

```{r}
# Ljung-Box test
sunspot.year %>%
  arima(c(9,0,0)) %>%
  diag_ljungbox %>%
  autoplot

# Box-Pierce Test
sunspot.year %>%
  arima(c(9,0,0)) %>%
  diag_boxpierce %>%
  autoplot
```

## Residual vs. Fitted Values (Developmental)

**Warning: The inputs for this function will change.**

One further useful plot is to compare the residuals to the fitted values.

```{r, fig.width = 5}
xt = sunspot.year
model = arima(xt, c(9,0,0))

resids = resid(model)
fit = xt - resids

diag_fitted(fit, resids) %>%
  autoplot
```

## Master Diagnostic Plot (Developmental)


**Warning: This function will change in the future to adhere more to a chained
argument**

Each of the diagnostic plots are available standalone. However, there is benefits
to plotting all of the diagnostic information simultaneously. Hence, the `diag_ts()`
function that will calculate all of the above values and provide a diagnostic plot.

```{r, fig.width = 5}
xt = sunspot.year
model = arima(xt, c(9,0,0))
diag_ts(model, xt) %>% 
  autoplot
```

# Selecting an appropriate time series model

There are many ways to go about selecting an ideal time series model. Within
the package there are `select_arima()`, `select_arma()`, `select_ar()`, 
and `select_ma()` functions that build all of the time series models alongside
multiple Information Criterion (IC) such as:

- Akaike information criterion (AIC): $AIC = - 2{L_{\max }} + 2k$
- Bayesian information criterion (BIC): $BIC = - 2{L_{\max }} + klog(n)$
- Hannan–Quinn information (HQC): $HQC =  - 2{L_{\max }} + 2k\log \left( {\log \left( n \right)} \right)$. 

where $L_{\max }$ is the log-likelihood, $k$ is the number of parameters, 
and $n$ is sample size.

To acess the best model according to one of the criteria, use: `best_model(x, ic="aic")`

## Selecting one dimension `select_ar()` and `select_ma()`

```{r}
# Bring in time series data
sunspot.year %>%
  gts(start = 1700, freq = 1) %>%
  select_ar(p.min = 1, p.max = 8) -> model_info

# Obtain a plot of the model selection criteria
model_info %>% 
  autoplot

# Obtain the best model (AIC is default)
model_info %>%
  best_model

# Obtain the best model with BIC
model_info %>%
  best_model(ic = 'bic') -> bmodel

# See the ACF of the residuals for the model
bmodel %>%
  emp_corr %>%
  autoplot
```

# Utilities

The package hosts various utilities to improve working with simulated data.

In particular, the package offers the ability to cast a simulation `matrix` to
`data.frame` for graphing within `ggplot2`. The simulation cast feature is
meant to ease the use of simulated data within `ggplot2`. Thus, support exists
for both casting simulated data stored by row (default) and by column into a
`Round` and `Draw` framework, where `Round` means the simulation iteration and
`Draw` means the statistic observed.

```{r}
# Set Seed
set.seed(5812)

# Generate data
m = matrix(rnorm(10), 2, 5)

# Organize by row
m %>%
  cast_simdf %>%
  head(n = 5)

# Organize by column
m %>%
  cast_simdf(wide = FALSE) %>%
  head(n = 5)

# Graph
m %>%
  cast_simdf(wide = FALSE) %>%
  autoplot
```
